Discovering causal factors with back-and-forth regression

### 2019-05-16 paper/paper.tex
break lines
fix stuff

\begin{abstract}
Identifying causes from observations is at the core of science. This endeavor is particularly challenging when i) potential factors are difficult to manipulate individually and ii) observations are complex and multi-dimensional. To address this issue, we introduce ``back-and-forth regression'' (BFR), a method that optimally identifies the most plausible causal factors from a set of covarying candidates. BFR consists in i) predicting the potential causes $X$ from the observable $Y$, and subsequently ii) identifying the only factors that contributed to those predictions. After detailing the proof of convergence, we show that BFR outperforms least-squares and related regression techniques, as well as cross-decomposition methods (e.g. canonical correlation analysis, and partial least squares) on two causal identification and prediction out of distribution. We apply BFR to several empirical datasets to show that it retrieves plausible results. In particular, we apply BFR to neuroimaging recordings of 102 subjects presented with word sequences, and show, as expected, that the early brain responses appear to be specifically caused by low-level visual factors whereas late brain responses appear to be caused lexico-semantic factors, despite the fact that these factors covary. We finish extending BFR to nonlinear cause-effect relationships.
\end{abstract}

>>>>

\begin{abstract}
Identifying causes from observations is at the core of science. This endeavor
is particularly challenging when i) potential factors are difficult to
manipulate individually and ii) observations are complex and multi-dimensional.
To address this issue, we introduce ``back-and-forth regression'' (BFR), a
method that optimally identifies the most plausible causal factors from a set
of covarying candidates. BFR consists of i) predicting the potential causes $X$
from the observable $Y$, and subsequently ii) identifying the only factors that
contributed to those predictions. After detailing the proof of convergence, we
show that BFR outperforms least-squares and related regression techniques, as
well as cross-decomposition methods (e.g. canonical correlation analysis, and
partial least squares) on two tasks: causal identification and out-of-distribution
prediction. We apply BFR to several empirical datasets to show that it
retrieves plausible results. In particular, we apply BFR to neuroimaging
recordings of 102 subjects presented with word sequences, and show, as
expected, that the early brain responses appear to be specifically caused by
low-level visual factors whereas late brain responses appear to be caused by
lexico-semantic factors, despite the fact that these factors covary. We finish
by extending BFR to nonlinear cause-effect relationships.
\end{abstract}
